---
title: "TNC_Analogs1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# This script processes and visualizes the effects of distance buffer, 
# climate bins and number of analogs on accuracy and agreement. 

# libraries
library(tidyverse)
library(splitstackshape)
library(stringr)
library(ggplot2)
library(patchwork)

# load the main Oregon climate data to look at the untransformed histograms of AET and Def: 

data<-readRDS("../outputs/XY.rds")
data0<-sample_n(data, size =10000, replace = FALSE)

# data where the number of analogs varies, but distance stays the same
n5<-read_csv("../outputs/accuracy_sqrt_05_125_50km_5analogs.csv")
n7<-read_csv("../outputs/accuracy_sqrt_075_125_50km_7analogs.csv")
n9<-read_csv("../outputs/accuracy_sqrt_075_125_50km_9analogs.csv")
n15<-read_csv("../outputs/accuracy_sqrt_075_125_50km_15analogs.csv")
n100<-read_csv("../outputs/accuracy_sqrt_075_125_50km_100analogs.csv")
n7_extra<-read_csv("../outputs/accuracy_sqrt_extra_bins7nn2.csv")
# remove 0.5 bins from n5 for now: 
n5.1<-n5[!grepl("0.5", n5$model), ]

```

# What do raw and sqrt transformed data look like:  

## Raw AET and deficit histograms: 

 * Data plotted are a random 10,000 pixel subsample of the 35X10^6 Oregon pixels (full dataset takes too long to plot)
 * These are "overlapping" histograms: histogram of non-forest plotted over histogram of forest pixels
 * Histogram bars are transparent to facilitate interpretation. 
```{r histograms, echo=FALSE, message=FALSE, warning=FALSE}


def_f_nf<-ggplot(data0, aes(x=def_h))+
  geom_histogram(data=subset(data0, forest==1), aes(fill="green"), alpha=0.2)+
  geom_histogram(data=subset(data0, forest==0), aes(fill="red"), alpha=0.2)+
  labs(title = "Raw Deficit Distribution (10 000 pixel subsample) ", color="Non-forest/Forest")+
  scale_fill_manual(name="Forest/non-forest", labels = c("red"="Non-forest", "green"="Forest"),   values=c("red"="red", "green"="green"))+
  xlab("Deficit")+
  theme_bw()
# 


aet_f_nf<-ggplot(data0, aes(x=aet_h))+
  geom_histogram(data=subset(data0, forest==1), aes(fill="green"), alpha=0.2)+
  geom_histogram(data=subset(data0, forest==0), aes(fill="red"), alpha=0.2)+
  labs(title = "Raw AET Distribution ", color="Non-forest/Forest")+
  scale_fill_manual(name="Forest/non-forest", labels = c("red"="Non-forest", "green"="Forest"),   values=c("red"="red", "green"="green"))+

  xlab("AET")+
  theme_bw()

# fig = ggplot(test, aes(x = value, fill = condition)) +
#     geom_histogram(position = "identity", alpha = .8) + 
#     scale_fill_manual(values=c("grey20", "grey60")) + theme_bw()

# (def/aet)
(def_f_nf/aet_f_nf)
```

## SQRT-transformed AET and deficit histograms: 


```{r sqare root histograms,  echo=FALSE, message=FALSE, warning=FALSE}

# for comparison: 
# def_f_nfsq<-ggplot(data0, aes(x=sqrt(def_h), color=as.factor(forest)))+
#   geom_histogram(fill="transparent")+
#   labs(title = "SQRT Deficit Distribution ", color="Non-forest/Forest")+
#   xlab("Deficit")+
#   theme_bw()


def_f_nfsq<-ggplot(data0, aes(x=sqrt(def_h)))+
  # geom_histogram()+
  geom_histogram(data=subset(data0, forest== 1), aes( fill="green"), alpha=0.2)+
  geom_histogram(data=subset(data0, forest== 0), aes( fill="red"), alpha=0.2)+

  # geom_histogram(subset(data0, forest == 0), aes(x=sqrt(aet_h)), alpha=0.2, fill="red")+
  # # geom_histogram(fill="transparent")+
  labs(title = "SQRT Deficit Distribution ", color="Non-forest/Forest")+
  # scale_color_manual(name = "Forest/non-forest", labels = c("Non-forest", "Forest"), values=c("red", "darkgreen")) +
    scale_fill_manual(name="Forest/non-forest", labels = c("red"="Non-forest", "green"="Forest"), values=c("red"="red", "green"="green"))+
  xlab("Deficit")+
  theme_bw()

aet_f_nfsq<-ggplot(data0, aes(x=sqrt(aet_h)))+
  # geom_histogram()+
  geom_histogram(data=subset(data0, forest== 1), aes( fill="green"), alpha=0.2)+
  geom_histogram(data=subset(data0, forest== 0), aes( fill="red"), alpha=0.2)+
  scale_fill_manual(name="Forest/non-forest", labels = c("red"="Non-forest", "green"="Forest"), values=c("red"="red", "green"="green"))+

  # geom_histogram(subset(data0, forest == 0), aes(x=sqrt(aet_h)), alpha=0.2, fill="red")+
  # # geom_histogram(fill="transparent")+
  labs(title = "SQRT AET Distribution ", color="Non-forest/Forest")+
  xlab("AET")+
  theme_bw()


# (def/aet)
(def_f_nfsq/aet_f_nfsq)
```


# What is the distribution of forests and non-forests in the AET/deficit climate space?

## 1= forest, 0 = non-forest.
```{r forest/non forest in the aet and def climate space, echo=FALSE,  message=FALSE, warning=FALSE}
data0$forest<-as.factor(data0$forest)
separate<-ggplot(data0, aes(x=def_h, y=aet_h))+
  geom_point()+
  theme_bw()+
  facet_wrap(vars(forest))+
  labs(title = "Forst & non-forest pixels in the climate space (apart)")+
  xlab("Deficit")+
  ylab("AET")


##############################################
together<-ggplot(data0, aes(x=def_h, y=aet_h, color= forest)) +
  geom_point(shape=1, alpha=0.5) +
 # geom_point(data=subset(data0, forest==1), colour="red", shape=1)+
    labs(title = "Non-forest & forest pixels in the climate space (together)")+
  xlab("Deficit")+
  ylab("AET")+
  theme_bw()


separate


together
```
 
 
 
 - ~ 400 mm deficit value seems like the the forest cutoff (in agreement with Stephenson 1990)
 - Forest and non-forest pixels share much of the climate space defined by AET and Deficit
 
 
 
# Where do the misclassified pixels lie in the climate space:
  * Using ~ 1 unit climate bin size with square-root transformed data
  * 7 analogs used for classification
  
  

```{r mislassification vis, echo = FALSE, message = FALSE, warning = FALSE}

# need to load the unprocessed classification data:
# do it just for the 7 analogs
mis<-readRDS("../outputs/accuracy_sqrt_circa_1mm_bins_7nn.rds")

# pick out misclassificaitons for one column only:
mis1<-mis%>%
  filter(forest.pred0.75_1!=forest)

# scatterplot of mistakes with the background of general forest/non-forest distribution

pt_plot<-ggplot(mis, aes(x=(def_h)^2, y=(aet_h)^2, color= as.factor(forest))) +
   geom_point(shape=1, alpha=0.5) +
  geom_point(data=mis1, aes(x=(def_h)^2, y=(aet_h)^2, color="misclassification"), shape =10)+
  theme_classic()+
  labs(title="Misclassified pixels in the AET & deficit space", subtitle ="1000 test pixels", color="Non-forest/Forest")+
  ylab("AET")+xlab("Deficit")

pt_plot
```



## Histogram of misclassificaitons by Defiict: 

```{r misclassification histograms, message=FALSE, warning=FALSE, echo=FALSE}

def_mistakes<-ggplot(data=data0, aes(x=def_h, color=as.factor(forest)))+
  geom_histogram(fill="transparent")+
  theme(legend.position = none)+
  geom_histogram(data=mis1, aes(x=(def_h)^2))+
  labs(title="Forest/non-forest histogram (hollow) and misclassified (solid) pixels by deficit", subtitle="Hollow bars - actual values from 10,000 pixel subsample. Solid bars = misclassifications", x="Deficit", color="misclassifications")+
  scale_color_manual(labels = c("Non-forest", "Forest"), values=c("red", "darkgreen")) +

  # ylim(0, 15)+
  theme_bw()

def_mistakes

# Zoomed-in histogram of mistakes by deficit: 
def_mistakes_z<-ggplot(data=mis1, aes(x=(def_h)^2))+
  geom_histogram(data=subset(mis1,forest.pred0.75_1==1), aes(fill="green"), alpha=0.2)+
  geom_histogram(data=subset(mis1,forest.pred0.75_1==0), aes(fill="red"), alpha=0.2)+
  labs(title="Misclassifications by deficit values (zoomed-in Y axix)", color="Misclassified as:")+
  scale_fill_manual(name="misclassifications", labels = c("red"="Non-forest", "green"="Forest"), values=c("red"="red", "green"="green"))+
  xlab("Deficit")+
  theme_bw()


def_mistakes_z

```


* Misclassifications are not limited to the vicinity of the 400 mm of deficit
* They happen at all the overlapping values of deficit

## Misclassifications by AET values:

```{r Misclassificaitons by AET just in case, message = FALSE, echo=FALSE, warning=FALSE}

# Misclassification by AET, just in case: 
aet_mistakes<-ggplot(data=data0, aes(x=aet_h, color=as.factor(forest)))+
  geom_histogram(fill="transparent")+
  theme(legend.position = none)+
  geom_histogram(data=mis1, aes(x=(def_h)^2))+
  labs(title="Forest/non-forest histogram and misclassified (solid) pixels by AET", subtitle="Hollow bars - actual values from 10,000 pixel subsample. Solid bars = misclassifications", x="AET", color="misclassifications")+
  scale_color_manual(labels = c("Non-forest", "Forest"), values=c("red", "darkgreen")) +

  # ylim(0, 15)+
  theme_bw()

aet_mistakes

# zoomed in view: 
aet_mistakes_z<-ggplot(data=mis1, aes(x=(aet_h)^2))+
 geom_histogram(data=subset(mis1,forest.pred0.75_1==1), aes(fill="green"), alpha=0.2)+
  geom_histogram(data=subset(mis1,forest.pred0.75_1==0), aes(fill="red"), alpha=0.2)+
  labs(title="Misclassifications by AET values (zoomed-in Y axix)", color="Misclassified as:")+
  scale_fill_manual(name="misclassifications", labels = c("red"="Non-forest", "green"="Forest"), values=c("red"="red", "green"="green"))+
  xlab("AET")+
  theme_bw()

aet_mistakes_z



```


<!-- # What is the effect of AET  and deficit window size on forest/non-forest classificaiton accuracy (sqrt-transformed data): -->
<!--  - These are results that I presented during our last meeting, but organized by AET and deficit values.  -->

<!--  - Multiple points within the analog number and deficit category represent different values of AET tested.  -->

<!--  - Multiple points within the analog number and AET category represent different values of deficit tested.  -->

<!--  - Is there a consistent effect of Deficit or AET bin size?  -->

<!--  - Smaller def bins are associated with higher accuracy (down to 0.75, but not lower)? but the effect is ephemeral, specific to N=7 -->
```{r influence of def values vs aet values: data, include = FALSE}
# head(n7)
# n7$n<-7
# n7_extra$n<-7
# n9$n<-9
# n15$n<-15
# n100$n<-100
# n5.1$n<-5
# 
# big_table<-rbind(n5.1,  n7[, -c(3:4)], n7_extra[, -c(3:4)], n9, n15[, -c(3:4)], n100) ### removing CI columns
# 
# str(big_table)
# 
# # separate out the bin values from "forest"
# big_table1<-big_table %>% 
#   separate(model, sep=11, into=c("forest", "new_model")) %>% 
#   separate(new_model, sep="_", into=c("def", "aet"), remove=FALSE)
# 
# 
# # convert model names to a factor: 
# big_table1$new_model<-as.factor(big_table1$new_model)
# 
# t_label<-as.character(big_table1$new_model)
# big_table1<-big_table1 %>% 
#   mutate(def=as.numeric(def))
# ```
# 
# 
# ```{r influence of def values vs aet values, echo=FALSE,  message=FALSE, warning=FALSE}
# # sort results by deficit values: 
# def<-big_table1 %>% 
#   arrange(by= "def") %>% 
#   ggplot( aes(x=def, y=accuracy, color=as.factor(n)))+
#   geom_point() +
#   theme_bw()+
#   labs(title = "Accuracy by deficit", color="# of analogs")+
#   xlab("Deficit bin size (sqrt(mm))")
#   
#   # scale_x_discrete(name="climate bins combinations (def_aet)", labels=t_label)+
#   # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# # sort results by aet values: 
# aet<-big_table1 %>% 
#   # filter(n=7)%>%
#   arrange(by= "aet") %>% 
#   ggplot( aes(x=aet, y=accuracy, color=as.factor(n)))+
#   geom_point() +
#   theme_bw()+
#   labs(title = "Accuracy by AET", color="# of analogs")+
#   xlab("AET bin size (sqrt(mm))")
# 
# (def/aet)
# 
# 
# ```
# ```{r forests & non-forests in the aet and def climate space,  message=FALSE, warning=FALSE}
# 
# 
# ```
# ```{r raw value tests that I already have, include=FALSE}
# 
# ## compile the wide bin tests that I already have: 
# raw0<-read_csv("../outputs/wide_bins_tests1.csv")
# 
# raw1<-read_csv("../outputs/medium_wide_bins_accuracy.csv")
# raw<-rbind(raw1[, c(2:5)], raw0)
# 
# ## separate out aet and deficit values: 
# 
# raw_t<-raw%>% 
#   separate(col="model", sep=11, into=c("model", "def_aet"), convert=TRUE) %>%
#   separate(col="def_aet", sep="_", into=c("def", "aet"), convert=TRUE)

```
# What is the effect of AET  and deficit window size on forest/non-forest classificaiton accuracy (raw data): 

* Re-ran tests for raw and square-root transformed data
* AET and deficit bin sizes are identical 
* Considered 7 analogs
* 50 km buffer 


```{r,  echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
# 
# 
# by_def<-ggplot(subset(raw_t, def<100), aes(x=def, y=Accuracy, color=as.factor(aet)))+
#   geom_point()+
#   scale_x_continuous(breaks=c(5, 10, 25, 50, 100, 250, 500))+
#   labs(title="Accuracy by Raw Def Bin Widths", color="AET bin size")+
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
#   theme_bw()+
#   xlab("Raw deficit bin size")
# 
# 
# 
# by_aet<-ggplot(subset(raw_t, aet<100), aes(x=aet, y=Accuracy, color=as.factor(def)))+
#   geom_point()+
#   scale_x_continuous(breaks=c(5, 10, 25, 50, 100, 250, 500))+
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
#   theme_bw()+
#   labs(title="Accuracy Organized by Raw AET Bin Widths", color="Deficit bin size")+
#   xlab("Raw AET bin size")
# 
# ### compare 
# 
# aet<-big_table1 %>% 
#   filter(n==7)%>%
#   arrange(by= "aet") %>% 
#   ggplot( aes(x=aet, y=accuracy, color=def))+
#   geom_point() +
#   theme_bw()+
#   labs(title = "accuracy by sqrt(aet) ", color="def")
#   
# 
# 
# def<-big_table1 %>% 
#   filter(n==7)%>%
#   arrange(by= "def") %>% 
#   ggplot( aes(x=def, y=accuracy, color=aet))+
#   geom_point() +
#   theme_bw()+
#   labs(title = "accuracy by sqrt(aet) ", color="aet")
#   
# 
# 
# (by_def/by_aet)
```
```{r,  message=FALSE, warning=FALSE, echo=FALSE}
# def
 
# aet

```

## Re-runs of the accuracy tests. Raw data: 
```{r echo=FALSE, message = FALSE}

test11<-read_csv("../outputs/accuracy_raw_reruns_5_50mm_7NN_50km.csv")
test12<-read_csv("../outputs/accuracy_raw_20_40mm_7NN_50km.csv")
test13<-read_csv("../outputs/accuracy_raw_reruns_75_100mm7NN.csv")
test14<-read_csv("../outputs/accuracy_raw_reruns_150mm_7NN_50km.csv")


new_raw<-rbind(test11, test12, test13, test14)



new_raw<-new_raw%>%
  separate(model, into=c("model", "def_aet"), sep=11)%>%
  separate(def_aet, into=c("def", "aet"), sep="_")

breaks<-as.numeric(unique(new_raw$aet))

ggplot(data=new_raw, aes(x=as.numeric(def), y=accuracy))+
  geom_point()+
  theme_bw()+
  ylim(0.7, 0.9)+
  labs(title = "Classificaition Accuracy, Raw Data, 7 Analogs")+
  scale_x_continuous(breaks=breaks)+
  xlab("Deficit & AET bin size (mm)")

```

## Re-runs of the accuracy tests, squre-root transformed data: 
```{r echo=FALSE, message = FALSE, warning = FALSE}

sqrt_reruns<-read_csv("../outputs/accuracy_sqrt_reruns_05_2_7nn_50km.csv")

# clean the data: 
new_sqrt<-sqrt_reruns%>%
  separate(model, into=c("model", "def_aet"), sep=11)%>%
  separate(def_aet, into=c("def", "aet"), sep="_")


# plot the data: 
breaks_sq<-as.numeric(unique(new_sqrt$def))

ggplot(data=new_sqrt, aes(x=as.numeric(def), y=accuracy))+
  geom_point()+
  theme_bw()+
  ylim(0.7, 0.9)+
  labs(title = "Classificaition Accuracy, SQRT Data, 7 Analogs")+
  scale_x_continuous(breaks=breaks_sq)+
  xlab("Deficit & AET bin size (sqrt(mm))")


```




# What is the effect of considering different number of analogs for projection classification? 


```{r echo=FALSE, message=FALSE, warning=FALSE}


# read in the data: 
nn5r<-read_csv("../outputs/accuracy_sqrt_reruns_05_2_5NN_50km.csv")
nn5r$n<-5
nn7r<-read_csv("../outputs/accuracy_sqrt_reruns_05_2_7nn_50km.csv")
nn7r$n<-7

nn9r<-read_csv("../outputs/accuracy_sqrt_reruns_05_2_9NN_50km.csv")
nn9r$n<-9
nn11r<-read_csv("../outputs/accuracy_sqrt_reruns_05_2_11NN_50km.csv")
nn11r$n<-11

nn15r<-read_csv("../outputs/accuracy_sqrt_reruns_05_2_15NN_50km.csv")
nn15r$n<-15


acc_nn<-rbind(nn5r, nn7r, nn9r, nn11r, nn15r)
acc_nn$n<-as.factor(acc_nn$n)


# clean: 
acc_nn1<-acc_nn%>%
  separate(model, into=c("model", "def_aet"), sep=11)%>%
  separate(def_aet, into=c("def", "aet"), sep="_")




ggplot(acc_nn1, aes(x=def, y=accuracy, color=n))+
  geom_point()+
  theme_bw()+
  labs(title= "Accuracy by number of analogs considered")+
  xlab("Deficit and AET Bin Size (sqrt(mm))")+
  ylim(0.7, 0.9)

### 
```

# Where are the 1000 focal plots located? 

```{r maps}
library(tidyverse)
library(raster)
aet<-raster("../data/aet_trouble/AET/AET.img")
data11<-readRDS("../outputs/accuracy2_2_11NN50km.rds")

# map misclassifications: 
mis<-readRDS("../outputs/accuracy_sqrt_circa_1mm_bins_7nn.rds")

# pick out misclassificaitons for one column only:
mis1<-mis%>%
  filter(forest.pred0.75_1!=forest)


plot(aet, main = "Location of one subset of 1000 focal pixels (black) and misclassified pixels (red)")
points(mis[, c(1:2)])
points(mis1[, c(1:2)], col='red', pch=12)

```

# Summary: 
## Effect of climate bin size:
- Raw and sqare-root transformed data appear to have equal accuracy.
    - Previous results, where raw data had lower accuracy were likely a result of a different search criteria - I looked at a number of them early on and did not label my results carefully. 

- Bin sizes (raw and square-root transformed) do not have a strong effect on accuracy. 

- Thoughts on the lack of bin size sensitivity:
  - Perhaps picking the "closest" analogs overrides the effect of widening the climate bin
  - Low sensitivity to bin sizes is in line with the broad overlap of forested and non-forested plots in the AET and Deficit space. Narrower bin is not necessarily meaningful. 


- The effect of # of analogs considered on accuracy is also weak, and is likely a spurious result of the focal subset considered (my algorithm picks a new set of 1000 focal plots for each test)

- Distance still has a strong effect on accuracy. 

