---
title: "TNC_Analogs1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# This script processes and visualizes the effects of distance buffer, 
# climate bins and number of analogs on accuracy and agreement. 

# libraries
library(tidyverse)
library(splitstackshape)
library(stringr)
library(ggplot2)
library(patchwork)

# load the main Oregon climate data to look at the untransformed histograms of AET and Def: 

data<-readRDS("../outputs/XY.rds")
data0<-sample_n(data, size =10000, replace = FALSE)

# data where the number of analogs varies, but distance stays the same
n5<-read_csv("../outputs/accuracy_sqrt_05_125_50km_5analogs.csv")
n7<-read_csv("../outputs/accuracy_sqrt_075_125_50km_7analogs.csv")
n9<-read_csv("../outputs/accuracy_sqrt_075_125_50km_9analogs.csv")
n15<-read_csv("../outputs/accuracy_sqrt_075_125_50km_15analogs.csv")
n100<-read_csv("../outputs/accuracy_sqrt_075_125_50km_100analogs.csv")
n7_extra<-read_csv("../outputs/accuracy_sqrt_extra_bins7nn2.csv")
# remove 0.5 bins from n5 for now: 
n5.1<-n5[!grepl("0.5", n5$model), ]

```

# What do raw and sqrt transformed data look like:  

## Raw AET and deficit histograms: 

```{r histograms, echo=FALSE, message=FALSE, warning=FALSE}
def_f_nf<-ggplot(data0, aes(x=def_h, color=as.factor(forest)))+
  geom_histogram()+
  labs(title = "Raw Deficit Distribution ", color="Non-forest/Forest")+
  xlab("Deficit")+
  theme_bw()

aet_f_nf<-ggplot(data0, aes(x=aet_h, color=as.factor(forest)))+
  geom_histogram()+
  labs(title = "Raw AET Distribution ", color="Non-forest/Forest")+
  xlab("AET")+
  theme_bw()


# (def/aet)
(def_f_nf/aet_f_nf)
```

## SQRT-transformed AET and deficit histograms: 
```{r sqare root histograms,  echo=FALSE, message=FALSE, warning=FALSE}

# for comparison: 
def_f_nfsq<-ggplot(data0, aes(x=sqrt(def_h), color=as.factor(forest)))+
  geom_histogram()+
  labs(title = "SQRT Deficit Distribution ", color="Non-forest/Forest")+
  xlab("Deficit")+
  theme_bw()

aet_f_nfsq<-ggplot(data0, aes(x=sqrt(aet_h), color=as.factor(forest)))+
  geom_histogram()+
  labs(title = "SQRT AET Distribution ", color="Non-forest/Forest")+
  xlab("AET")+
  theme_bw()


# (def/aet)
(def_f_nfsq/aet_f_nfsq)
```


# What is the distribution of forests and non-forests in the AET/deficit climate space?

## 1= forest, 0 = non-forest.
```{r forest/non forest in the aet and def climate space, echo=FALSE,  message=FALSE, warning=FALSE}
data0$forest<-as.factor(data0$forest)
separate<-ggplot(data0, aes(x=def_h, y=aet_h))+
  geom_point()+
  theme_bw()+
  facet_wrap(vars(forest))+
  labs(title = "Forst & non-forest pixels in the climate space (apart)")+
  xlab("Deficit")+
  ylab("AET")


##############################################
together<-ggplot(data0, aes(x=def_h, y=aet_h, color= forest)) +
  geom_point(shape=1, alpha=0.5) +
 # geom_point(data=subset(data0, forest==1), colour="red", shape=1)+
    labs(title = "Non-forest & forest pixels in the climate space (together)")+
  xlab("Deficit")+
  ylab("AET")+
  theme_bw()


separate


together
```
 
 
 
 - ~ 400 mm deficit value seems like the the forest cutoff (in agreement with Stephenson 1990)
 - Other climate thresholds?
 
# What is the effect of AET  and deficit window size on forest/non-forest classificaiton accuracy (sqrt-transformed data):
 - These are results that I presented during our last meeting, but organized by AET and deficit values. 

 - Multiple points within the analog number and deficit category represent different values of AET tested. 

 - Multiple points within the analog number and AET category represent different values of deficit tested. 
 
 - Is there a consistent effect of Deficit or AET bin size? 
 
 - Smaller def bins are associated with higher accuracy (down to 0.75, but not lower)? but the effect is ephemeral, specific to N=7
```{r influence of def values vs aet values: data, include = FALSE}
head(n7)
n7$n<-7
n7_extra$n<-7
n9$n<-9
n15$n<-15
n100$n<-100
n5.1$n<-5

big_table<-rbind(n5.1,  n7[, -c(3:4)], n7_extra[, -c(3:4)], n9, n15[, -c(3:4)], n100) ### removing CI columns

str(big_table)

# separate out the bin values from "forest"
big_table1<-big_table %>% 
  separate(model, sep=11, into=c("forest", "new_model")) %>% 
  separate(new_model, sep="_", into=c("def", "aet"), remove=FALSE)


# convert model names to a factor: 
big_table1$new_model<-as.factor(big_table1$new_model)

t_label<-as.character(big_table1$new_model)
big_table1<-big_table1 %>% 
  mutate(def=as.numeric(def))
```


```{r influence of def values vs aet values, echo=FALSE,  message=FALSE, warning=FALSE}
# sort results by deficit values: 
def<-big_table1 %>% 
  arrange(by= "def") %>% 
  ggplot( aes(x=def, y=accuracy, color=as.factor(n)))+
  geom_point() +
  theme_bw()+
  labs(title = "Accuracy by deficit", color="# of analogs")+
  xlab("Deficit bin size (sqrt(mm))")
  
  # scale_x_discrete(name="climate bins combinations (def_aet)", labels=t_label)+
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# sort results by aet values: 
aet<-big_table1 %>% 
  # filter(n=7)%>%
  arrange(by= "aet") %>% 
  ggplot( aes(x=aet, y=accuracy, color=as.factor(n)))+
  geom_point() +
  theme_bw()+
  labs(title = "Accuracy by AET", color="# of analogs")+
  xlab("AET bin size (sqrt(mm))")

(def/aet)


```
```{r forests & non-forests in the aet and def climate space,  message=FALSE, warning=FALSE}


```
```{r raw value tests that I already have, include=FALSE}

## compile the wide bin tests that I already have: 
raw0<-read_csv("../outputs/wide_bins_tests1.csv")

raw1<-read_csv("../outputs/medium_wide_bins_accuracy.csv")
raw<-rbind(raw1[, c(2:5)], raw0)

## separate out aet and deficit values: 

raw_t<-raw%>% 
  separate(col="model", sep=11, into=c("model", "def_aet"), convert=TRUE) %>%
  separate(col="def_aet", sep="_", into=c("def", "aet"), convert=TRUE)

```
# What is the effect of AET  and deficit window size on forest/non-forest classificaiton accuracy (raw data): 

```{r,  echo=FALSE, message=FALSE, warning=FALSE}
# 
# 
# by_def<-ggplot(subset(raw_t, def<100), aes(x=def, y=Accuracy, color=as.factor(aet)))+
#   geom_point()+
#   scale_x_continuous(breaks=c(5, 10, 25, 50, 100, 250, 500))+
#   labs(title="Accuracy by Raw Def Bin Widths", color="AET bin size")+
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
#   theme_bw()+
#   xlab("Raw deficit bin size")
# 
# 
# 
# by_aet<-ggplot(subset(raw_t, aet<100), aes(x=aet, y=Accuracy, color=as.factor(def)))+
#   geom_point()+
#   scale_x_continuous(breaks=c(5, 10, 25, 50, 100, 250, 500))+
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
#   theme_bw()+
#   labs(title="Accuracy Organized by Raw AET Bin Widths", color="Deficit bin size")+
#   xlab("Raw AET bin size")
# 
# ### compare 
# 
# aet<-big_table1 %>% 
#   filter(n==7)%>%
#   arrange(by= "aet") %>% 
#   ggplot( aes(x=aet, y=accuracy, color=def))+
#   geom_point() +
#   theme_bw()+
#   labs(title = "accuracy by sqrt(aet) ", color="def")
#   
# 
# 
# def<-big_table1 %>% 
#   filter(n==7)%>%
#   arrange(by= "def") %>% 
#   ggplot( aes(x=def, y=accuracy, color=aet))+
#   geom_point() +
#   theme_bw()+
#   labs(title = "accuracy by sqrt(aet) ", color="aet")
#   
# 
# 
# (by_def/by_aet)
```
```{r,  message=FALSE, warning=FALSE}
# def
 
# aet

```

```{r, echo=FALSE, warning=FALSE, message = FALSE}

test11<-read_csv("../outputs/accuracy_raw_reruns_5_50mm_7NN_50km.csv")
test12<-read_csv("../outputs/accuracy_raw_20_40mm_7NN_50km.csv")
test13<-read_csv("../outputs/accuracy_raw_reruns_75_100mm7NN.csv")
test14<-read_csv("../outputs/accuracy_raw_reruns_150mm_7NN_50km.csv")


new_raw<-rbind(test11, test12, test13, test14)


head(new_raw)
tail(new_raw)

new_raw<-new_raw%>%
  separate(model, into=c("model", "def_aet"), sep=11)%>%
  separate(def_aet, into=c("def", "aet"), sep="_")
head(new_raw)


ggplot(data=new_raw, aes(x=as.numeric(def), y=accuracy))+
  geom_point()+
  theme_bw()+
  ylim(0.7, 0.9)+
  labs(title = "Classificaition Accuracy, Raw Data, 7 Analogs")+
  xlab("Deficit & AET bin size (mm)")

```


<!-- # Where do the misclassified pixels lie in climate space: -->

<!-- ```{r mislassification vis, echo = FALSE, message = FALSE, warning = FALSE} -->

<!-- # need to load the unprocessed classification data: -->
<!-- # do it just for the 7 analogs -->
<!-- mis<-readRDS("../outputs/accuracy_sqrt_circa_1mm_bins_7nn.rds") -->

<!-- # pick out misclassificaitons for one column only: -->
<!-- mis1<-mis%>% -->
<!--   filter(forest.pred0.75_1!=forest) -->

<!-- # visualize, where these misclassifications are happening: -->
<!-- # ggplot(data0, aes(x=def_h, y=aet_h, color= forest)) + -->
<!-- #   geom_point(shape=1, alpha=0.5) + -->
<!-- #   geom_point(data=mis1, aes(x=(def_h)^2, y=(aet_h)^2, color=forest), -->
<!-- #              shape = 3) -->


<!-- # ggplot(data0, aes(x=def_h, y=aet_h, color= forest)) + -->
<!-- #   geom_point() + -->
<!-- #   geom_point(data=mis1, aes(x=(def_h)^2, y=(aet_h)^2, fill=as.factor(forest))) -->

<!-- pt_plot<-ggplot(mis, aes(x=(def_h)^2, y=(aet_h)^2, color= as.factor(forest))) + -->
<!--    geom_point(shape=1, alpha=0.5) + -->
<!--   geom_point(data=mis1, aes(x=(def_h)^2, y=(aet_h)^2, color="mistake"), shape =10) -->

<!-- # zoom<-ggplot(data=data0, aes(x=def_h, color=forest))+ -->
<!-- #   geom_histogram()+ -->
<!-- ggplot()+ -->
<!--   geom_histogram(data=mis1, aes(x=(def_h)^2, color=as.factor(forest)))+ -->
<!--   ylim(0, 10)+ -->
<!--     labs(color="incorrectly classified as ")+ -->
<!--   theme_bw() -->

<!-- def_mistakes<-ggplot(data=data0, aes(x=def_h, color=forest))+ -->
<!--   geom_histogram()+ -->
<!--   geom_histogram(data=mis1, aes(x=(def_h)^2, color='mistake'), fill='yellow')+ -->
<!--   theme_bw() -->

<!-- # geom_point(data=mis1, aes(x=(def_h)^2, y=(aet_h)^2, color=as.factor(forest)), shape =10) -->

<!-- (zoom+def_mistakes) -->

<!-- aet_mistakes<-ggplot(data=data0, aes(x=aet_h, color=forest))+ -->
<!--   geom_histogram()+ -->
<!--   geom_histogram(data=mis1, aes(x=(aet_h)^2, color='mistake'), fill='yellow')+ -->
<!--   theme_bw() -->

<!-- aet_m_zoom<-ggplot(data=data0, aes(x=aet_h, color=forest))+ -->
<!--   geom_histogram()+ -->
<!--   geom_histogram(data=mis1, aes(x=(aet_h)^2, color='mistake'), fill='yellow')+ -->
<!--   ylim(0, 10)+ -->
<!--   theme_bw() -->


<!-- (aet_mistakes+ aet_m_zoom) -->
<!-- ``` -->



## Summary: 
- Using square-root transformed data results in higher overall classification accuracy compared to raw data.

- Perhaps the result of the "dynamic" climate bin size: smaller bins for lower values of AET and Deficit, 
larger bins for high values of AET and Deficit. 
